MESOSCOPE manuscript

### 2023-09-06 manuscript/manuscript.Rmd
Adding Methods section from masters manuscript

# Methods
# Conclusions
# References

>>>>

Despite its historical reputation as a static and unchanging environment, the North Pacific Subtropical Gyre shows significant variation in its particulate composition due to mesoscale eddies.

# Methods

## Cruise information
Samples were collected from two cruises in the North Pacific Subtropical Gyre near Station ALOHA that targeted strong mesoscale eddy features as described in @Dugenne2023 and @Gleich2023. Briefly, the 2017 MESOSCOPE cruise (KM1709 on the *R/V Kilo Moana*) consisted of a transect across the cyclonic and anticyclonic poles of the eddy dipole as well as two long-term Lagrangian stations at the center of each eddy. The cyclonic pole of the eddy had a maximum negative SLA anomaly of -20 centimeters and the anticyclonic center reached +24 centimeters. The transect samples were taken at various times of day as the ship transected the dipole, and the eddy center samples were all collected between 5 and 8 pm. In 2018, the Falkor cruise (FK180310 on the *R/V Falkor*) targeted a new eddy dipole in approximately the same location with samples taken at the center of each eddy (maximum negative anomaly in the cyclone = -15 cm, maximum positive anomaly in the anticyclone = 26 cm).

## Biogeochemical data

Biogeochemical measurements were collected as described in @Barone2022 and @Dugenne2023 mostly following protocols used by the HOT program (http://hahana.soest.hawaii.edu/hot/methods/results.html). Briefly, all environmental data was collected via CTD Niskin bottles except beam attenuation which was measured via Hyperpro and sea level anomaly (SLA) which was measured via satellite. Chlorophyll and pheophytin were measured fluorometrically, particulate carbon (PC) and particulate nitrogen (PN) were measured using an elemental analyzer, and nitrate + nitrite (N+N) and soluble reactive phosphorus (phosphate, PO$_4^{3-}$) were measured on an autoanalyzer except where concentrations were below 100nM in which case they were measured using chemiluminescence. Values were interpolated linearly to the locations at which metabolite samples were taken using all data collected at the given station.

## Metabolite sample collection
Samples were obtained using the onboard CTD rosette to collect water from 15 meters, the deep chlorophyll maximum (DCM), and 175 meters during the eddy transect and from the DCM ±10 and ±20 meters during the eddy center sampling in the MESOSCOPE cruise and from 25 meters and the DCM during the Falkor cruise. The DCM was determined visually from fluorometer data during the CTD downcast and Niskin bottles were tripped during the return trip to the surface. Seawater from each depth was sampled in triplicate by firing one Niskin bottle for each sample. Samples were brought to the surface and decanted into prewashed (3x with DI, 3x with sampled seawater) polycarbonate bottles for filtration. Samples were filtered by peristaltic pump onto 142mm 0.2 µm Durapore filters held by polycarbonate filter holders on a Masterflex tubing line. Pressures were kept as low as possible while still producing a reasonable rate of flow through the filter, approximately 250-500 mL per minute. Samples were then removed from the filter holder using solvent-washed tweezers and placed into pre-combusted aluminum foil packets that were then flash-frozen in liquid nitrogen before being stored at -80 °C until extraction. A methodological blank was also collected by running filtrate through a new filter and then treated identically to the samples.

## Metabolite sample extraction
Extraction followed a modified Bligh & Dyer approach as detailed in @Boysen2018. Briefly, filters were added to PTFE centrifuge tubes with a 1:1 mix of 100 µm and 400 µm silica beads, approximately 2mL -20 °C Optima-grade DCM, and approximately 3mL -20 °C 1:1 methanol/water solution (both also Optima-grade). Extraction standards were added during this step. The samples were then bead-beaten three times, followed by triplicate washes with fresh methanol/water mixture. Samples were then dried down under clean nitrogen gas and warmed using a Fisher-Scientific Reacti-Therm module. Dried aqueous fractions were re-dissolved in 380 µL of Optima-grade water and amended with 20 µL isotope-labeled injection standards. Additional internal standards were added at this point to measure the variability introduced by chromatography and ionization, and the reconstituted fraction was syringe-filtered to remove any potential clogging material. This aqueous fraction was then aliquoted into an HPLC vial for injection on the HILIC column and diluted 1:1 with Optima-grade water. A pooled sample was created by combining 20 µL of each sample into the same HPLC vial, and a 1:1 dilution with water half-strength sample was aliquot from that to assess matrix effects and obscuring variation [@Boysen2018]. Also run alongside the environmental samples were two mixes of authentic standards in water and in an aliquot of the pooled sample at a variety of concentrations for quality control, annotation, and absolute concentration calculations. HPLC vials containing the samples were frozen at -80 °C until thawing shortly before injection.

## HPLC-MS methods
Separate HPLC-MS runs were used for the MESOSCOPE eddy transect, the MESOSCOPE eddy center, and the Falkor cruise data. For each, a SeQuant ZIC-pHILIC column (5 um particle size, 2.1 mm x 150 mm, from Millipore) was used with 10 mM ammonium carbonate in 85:15 acetonitrile to water (Solvent A) and 10 mM ammonium carbonate in 85:15 water to acetonitrile (Solvent B) at a flow rate of 0.15 mL/min. The column was held at 100% A for 2 minutes, ramped to 64% B over 18 minutes, ramped to 100% B over 1 minute, held at 100% B for 5 minutes, and equilibrated at 100% A for 25 minutes (50 minutes total). The column was maintained at 30 °C. The injection volume was 2 µL for samples and standard mixes. When starting a batch, the column was equilibrated at the starting conditions for at least 30 minutes. To improve the performance of the HILIC column, we maintained the same injection volume, kept the instrument running water blanks between samples as necessary, and injected standards in a representative matrix (the pooled sample) in addition to standards in water. After each batch, the column was flushed with 10 mM ammonium carbonate in 85:15 water to acetonitrile for 20 to 30 minutes.

Metabolomic data was collected on a Thermo Q Exactive HF hybrid Orbitrap (QE) mass spectrometer. The capillary and auxiliary gas heater temperatures were maintained at 320°C and 100°C, respectively. The S-lens RF level was kept at 65, the H-ESI voltage was set to 3.3 kV and sheath gas, auxiliary gas, and sweep gas flow rates were set at 16, 3, and 1, respectively. Polarity switching was used with a scan range of 60 to 900 *m/z* and a resolution of 60,000. Calibration was performed every 3-4 days at a target mass of 200 *m/z*. DDA data was collected from the pooled samples for high-confidence annotation of knowns and unknowns. All files were then converted to an open-source mzML format and centroided via Proteowizard’s msConvert tool.

## Targeted metabolomic methods

Targeted compounds were manually integrated using Skyline and MSDIAL. Skyline was used for both the eddy center samples and the Falkor cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset. Raw peak areas were then normalized to their best-matched internal standard per @Boysen2018 except that all compounds were normalized to their best match no matter how much or how little improvement was detected. Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standards and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.

## Untargeted metabolomic methods

Samples were also processed via the untargeted workflow detailed in @Kumler2023. In short, XCMS was used for peak detection and followed by manual inspection of each mass feature for quality with only those peaks classified as "Good" used in the analysis here. Features were then matched with an internal standard as described above.

## Statistics

All statistics were run in R version `r paste(version$major, version$minor, sep=".")`. We used multivariate statistics provided by the vegan package [@R-vegan] to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses. Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible. Multiple comparison problems were controlled where necessary using a false-discovery rate correction.

# Conclusions
# References
